---
title: "KidneyDonationAnalysis"
author: "Anjay Friedman"
date: "11/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
getwd()
setwd("/Users/anjayfriedman1/Downloads")
kidneyData <- read.csv("Seth_Kidney_Data.csv", header = TRUE)
library(ggplot2)


```

```{r}
head(kidneyData)
labels(kidneyData)
dim(kidneyData)


#explore how population relates to OPO.obsered.donation.rate.per.100.eligible.deaths

ggplot(kidneyData, aes(x = Population, y = OPO.observed.donation.rate.per.100.eligible.deaths)) + geom_point()

#model 
popVersusRate <- lm( OPO.observed.donation.rate.per.100.eligible.deaths ~ Population , data = kidneyData)
#formula: Rate per 100 = C + B*Population //equation we are estimating
summary(popVersusRate)
#R-squared is very low, so low chance of them being related 

```

Conclusion: It is not likely that there is a relationship between the population and donatation rate per 100 


New Question: How does income affect organ donations?
```{r}
labels(kidneyData)
```


Other Question: How does patients on kidney list affect donations?
```{r}
#patients on kidney list versus donors per million
ggplot(kidneyData, aes(y = Donors.per.million.population, x = Patients.on.list...kidney)) + geom_point()

#model 
listSizeVersusDperMillion <- lm( Donors.per.million.population ~ Patients.on.list...kidney , data = kidneyData)
#formula: Rate per 100 = C + B*Population //equation we are estimating
summary(listSizeVersusDperMillion)

#patients on kidney list versus donors per 10,000 deaths
ggplot(kidneyData, aes(y = Donors.per.10.000.deaths, x = Patients.on.list...kidney)) + geom_point()

#model 
listSizeVersusDper10 <- lm( Donors.per.10.000.deaths ~ Patients.on.list...kidney , data = kidneyData)
#formula: Rate per 100 = C + B*Population //equation we are estimating
summary(listSizeVersusDper10)
#R-squared is very low, so low chance of them being related 
```

Both have very low R-squared values --> It appears they are not linked.
This is interesting. I thought that if there were more people on the list, there would have been more donors due to:
a. Increased demand, therefore, increased pressure to donate and increased awareness
b. More people on list means more potential family donations 

